Integrating brain and biomechanical models - a new paradigm for understanding neuro-muscular control

### 2017-12-08 papers/integration/omm_integ.tex
Minor fixes, following proof-read of manuscript

% Solution described means we now need to do background of eye movements:
In an attempt to build a model combining brain, realistic
biomechanics \e{and} sensory feedback, we sought to extend our
\rvw{previous work modelling the oculomotor system by adding a virtual,
biomechanical eye model able to make physically realistic
movements. The rotational state of the eye would then determine how
visual features in the virtual world were projected back into the
brain model.}
%
The existing
model \citep{cope_basal_2017} is already able to capture sensory input
and convert it into a neural signal, assumed to specify the target of
a \e{saccadic eye movement}; a fast movement of the eyes which directs
the fovea to a region of interest in the field of view.
%
The oculomotor system is an excellent candidate for modelling because
its movements can be specified with only three degrees of freedom,
making it one of the simplest neuro-muscular systems in the body.
%
It is nevertheless behaviourally interesting, as saccadic eye
movements reveal information about decision making at a subconscious
level
\citep{deubel_saccade_1996,reppert_modulation_2015,marcos_determining_2016}.
The modelling of the oculomotor system is served by a large body of
behavioural data describing
saccades \citep{tipper_reaching_2001,walker_effect_1997,casteau_effect_2012},
many anatomical studies of the neural substrates involved
\citep{meredith_intrinsic_1998,isa_intrinsic_2002,isa_exploring_2009}
and electrophysiological data linking these together
\citep{hepp_spatio-temporal_1983,dorris_neuronal_1997,mcpeek_competition_2003,vokoun_circuit_2011}.
Furthermore, in the context of building \e{behaving} systems, a
necessary part of any model for which the behaviour requires visual
attention and decision making is a realistic mechanism for gathering
visual information.  This is obvious from extrinsic considerations---a
subject must look at a scene to make decisions or navigate within
it. It also follows for \e{intrinsic} reasons. For
example, \cite{howard_hand_1997} showed that visual cues affect reach
trajectories and the same group later demonstrated that reaching
affects the saccadic system
\citep{tipper_reaching_2001} suggesting a close relationship between these
neural systems. \rvw{Building a behaving oculomotor system will
therefore assist future computational neurobehavioural modelling
efforts that involve reaching.}

Many neural populations are involved in the coding of saccadic eye
movements, only a very brief overview is given here; for a review,
see \cite{munoz_commentary:_2002}. One pathway takes information from
the retina directly into the superficial layers of the superior
colliculus in the brainstem
\citep{sterling_receptive_1971,linden_massive_1983,wu_involvement_1994}.
Activity within the superior colliculus then excites neurons in the
pons, medulla and rostral mid-brain \citep{sparks_brainstem_2002}.
and \rvw{finally the motor neurons} which innervate the extraocular muscles
\citep{fuchs_firing_1970,sparks_brainstem_2002}. This direct pathway
is responsible for the low latency saccades called express saccades
\citep{schiller_effect_1987,edelman_activity_1996}.
%
Information from the retina is also processed by visual cortex which
feeds through to the frontal eye fields in which activity is related
to reflexive and voluntary
saccades \citep{schall_neural_1999}. Activity build-up in the frontal
eye fields is transferred to the intermediate layers of the superior
colliculus
\citep{stanton_frontal_1988-1} and is also processed by the basal ganglia,
which participates in the selection of the winning saccade end point
\citep{stanton_frontal_1988,hikosaka_role_2000}.
%
Although both cortical and subcortical paths produce a saccade target
signal in the superior colliculus, it is also possible for animals to
make relatively normal saccades even after the colliculus has been
ablated
\citep{wurtz_activity_1972,aizawa_reversible_1998},
though express saccades are lost with collicular lesions
\citep{schiller_effect_1987}. This makes the superior colliculus
a perplexing structure, being both critically involved in saccade
target specification \citep{sparks_sensory_1987} and saccade dynamic control
\citep{waitzman_superior_1991,goossens_optimal_2012} and yet dispensible.
The `backup pathway' likely incorporates the oculomotor vermis and
fastigial oculomotor region of the cerebellum which are known to
participate in the specification, dynamics and adaptation of saccadic
eye movements \citep{kleine_saccade-related_2003,takagi_effects_1998}.

@@

% A very short introduction to the section.
The integrated brain and biomechanical model described here is a
development of the model in \cite{cope_basal_2017}, referred to here
as the \ccg~model. This was a rate-coded neural network model
incorporating retinal populations, frontal eye fields (FEF), the basal
ganglia (BG), and the superior colliculus (SC).  The \ccg~model takes
as \e{input} the positions of luminances \rvwt{(of fixed shape and
intensity)} on a topographic map. \rvwt{Whilst certain assumptions
were made about the input---that a luminant input excites activity on
a retinotopic layer, with computer code carrying out the
transformation achieved in the brain by a neural connectivity
map \citep{thivierge_topographic_2007}---it is nonetheless \e{not} an
input-assumption model according to our definition because the
activity generated in the neural input layer is modelled as a response
to the luminances, rather than being crafted.}  In the \ccg~model, the
centroid of the activity in the deep layers of superior colliculus was
assumed to accurately encode the location of the eye at the end of the
saccade
\citep{wurtz_activity_1972,robinson_eye_1972,van_gisbergen_collicular_1987,mcilwain_lateral_1982}.  This location was used to
recalculate the positions of the luminances in the eye's frame of
reference at each time step. \rvwt{Because a pattern of neural
activity in the output population was assumed to have a behavioural
outcome, it was thus an \e{output-assumption model}}.  The model
included no brainstem populations other than superior colliculus, nor
a neuromuscular model.

To the \ccg~model, we added a brainstem model and a biomechanical eye
model. \rvw{The rate-coded brainstem model was taken from the
literature} \citep{gancarz_neural_1998} as the best-of-breed saccadic
burst generator \citep{girard_brainstem_2005}.  The biomechanical eye
was implemented using the biomechanical modelling framework OpenSim;
the brain and brainstem were modelled using the SpineML toolchain.
These will be described below, \rvwt{along with a review of
the \ccg~model}, but first we will give a description of the
co-ordinate systems.

@@

% Describe what is modelled first. This para is 'The OM system
% performs action selection and the BG will do that for us.'
The brain model, excluding the brainstem, is a re-implementation of
the \ccg~model, \rvwt{of reflexive saccadic behaviour \citep{cope_basal_2017}}.
\rvwt{Reflexive saccades are fast eye movements elicited by abrupt
changes in the peripheral visual scene (reflexive saccades can occur
also as a result of auditory and somatosensory stimuli, but these
modalities are ignored in this model). A reflexive saccade has a
starting position defined by the initial orientation of the eye and an
end-point position in which the eye is directed towards a new
target. Regardless of the number of targets within the visual scene,
the brain must choose one location as the end-point, because the eyes
can look only in one direction at a time. The functionality reproduced
by the \ccg~model is `the selection of the best target end-point for a
reflexive saccade'. A competition such as this between incompatible
movements is often referred to as an \e{action selection}
problem \citep{norman_attention_1986,maes_dynamics_1989,redgrave_basal_1999}.
The \ccg~model is therefore a model of action-selection in the
oculomotor system for reflexive saccades.  One hypothesis for the
r\^ole played by the basal ganglia (BG) is that the system
performs \e{action selection}
\citep{mink_basal_1996,redgrave_basal_1999,hikosaka_role_2000}.
The \ccg~model places the BG at the centre of the oculomotor system;
this follows the known anatomy of the
region \citep{hikosaka_role_2000} and provides a mechanism for action
selection of the best saccade. The BG receives input indirectly from
the superior colliculus, which has a retinotopic
arrangement \citep{ottes_visuomotor_1986}.}

@@

% Now describe each bit a little more, with help of the CCG
% macroscopic architecture figure.
\rvwt{Fig.~\ref{fig:ccg_model}(a) shows the macroscopic architecture of
the \ccg~model. The figure shows the relationships between the retinal
input populations, the FEF, the populations comprising the BG
sub-system (the red border indicates that the box represents a number
of populations as a sub-system), the thalamus and the superior
colliculus. Excitatory connections are indicated with arrow heads;
inhibitory connections with circles in place of the arrow heads. The
blue and green connection lines indicate two thalamo-basal ganglia
loops, one cortical loop through FEF (green), the other a sub-cortical
loop through SC. It is important to note that although they are given
different colours in the diagram, these loops are in no way
independent, with loop activity combining both in thalamus and in the
basal ganglia and a direct excitatory, feed-forward connection from
FEF to SC.}

@@

% Dopamine, tonic output.
\rvwt{At the neuronal level, the STN, GPe and SNr have tonic output
levels \citep{chevalier_disinhibition_1990,delong_primate_1985,kita_intracellular_1991}. This
is modelled using piecewise linear output functions with zero offsets,
$c$ (see Eq.~\ref{eq:lintransfunc}) but with noise added to the
input. In striatum, Str\_D1 and Str\_D2 have positive offset $c$,
mimicking the so-called `down-state' of medium spiny neurons which
have a resting potential far below spiking threshold and require
co-ordinated input to generate action
potentials \citep{wilson_origins_1996}.
% Note, in Alex's paper, there is an error with the offsets for D1,
% D2. These are set to -0.05 instead of 0.05. A quick trial with the
% model suggests this doesn't make a very large difference to the
% overall operation (you just see a background of noise around the
% signals in Str), and I have the sign correct in any case.
In addition, the Str\_D1 and Str\_D2 neurons are influenced by
dopamine in different ways; facilitating cortico-striatal transmission
at medium spiny neurons with D1
receptors \citep{hernandez-lopez_D1_1997,gonon_prolonged_1997} and
reducing transmission at those with D2
receptors \citep{delgado_inhibitory_1999}. These effects are modelled
using a dopamine parameter, $d$, which modulates the input activations
$a_{in}^D1$ and $a_{in}^D1$ as:
\begin{equation}
   a_{in}^{D1} = (0.2 + d)A
\end{equation}
\begin{equation}
   a_{in}^{D2} = (1 - d)A
\end{equation}
where $A$ is the input activation (see also
Eqns.~\ref{eq:a_in_d1} \& \ref{eq:a_in_d2}). For the `normal, healthy'
value for $d$ of 0.7, Str\_D1 activation is relatively enhanced
(0.9$A$); Str\_D2 activation is one third of this value (0.3$A$). The
major effect of this difference in the relative strength of the
activity in Str\_D1 versus Str\_D2 is simply that a change in the
level of activity in Str\_D1 affects the off-centre, on-surround
mechanism. The effect of varying the input into Str\_D2 is much more
subtle, with only a small change in the amount of inhibition fed from
GPe into STN (via a focussed, one to one connection) being affected by
the change, along with a small change in the inhibition fed into SNr
from GPe (also via a one to one connection). It is not possible to
clearly describe the dynamic effect of the dopamine parameter as being
due to any single population, because the activity is recurrently
connected through multiple loops. Thus, a line of reasoning such as
``reduced activity for a luminance in Str\_D2 will lead to less
inhibition in that region in GPe, which means that there will be
higher activity there, and hence more inhibition for that region
passed to STN leading us to expect less activity in STN'' is verified
by running a suitable simulation with the model, but the effect is
small. }
%
\rvwt{Note that
the effect of dopamine in the model is only to modulate the strength
of cortico-striatal synapses; no learning is modelled and so the
significance of dopamine as a prediction error signal is outside the
scope of the current work.}

% Change this - we already talked about channels in the context of the
% three channel model shown. Now talk about how we have a map of
% channels - a continuous space in the \ccg model. This can then lead
% on to Fig 3.
\rvwt{The GPR model in Fig.~\ref{fig:ccg_model}(b) has only three
channels, with the focussed inhibition from striatum to SNr and GPe
defined by a simple one-to-one scheme. The action channels represent
discrete, incompatible motor action choices. In the oculomotor model,
an action channel represents the end-point of saccade, and the
competition carried out in the basal ganglia is between potential
saccade end-points. However, eye movements have a \e{continuous}
end-point space; the eye can rotate to any orientation within its
biomechanically permissible range. Some end-points within this range
are mutually exclusive---it's not possible to look to the left and to
the right simultaneously---but \e{nearby} end-points are not
necessarily incompatible. A small enough error in the end-point of a
saccade will not prevent the eye from foveating on a target as the
foveal region of high visual acuity is not infinitesimally small. To
cope with this requirement, the populations within the oculomotor
basal ganglia are conceived of as two-dimensional topographic grids of
leaky integrator neural elements. Activity in each element corresponds
to a spatial location in the visual field. Neighbouring elements
correspond to locations which are close to each other in the visual
field. Focussed one-to-one projections in the GPR model are replaced
by projective fields with many weighted connections. Specifically,
each unit in Str\_D1 projected to a counterpart SNr$_j$ in SNr with
some weight $w_{max}$, but also connected to neighbouring nodes in SNr
with a weight given by $w_{max}.G(d)$, where $G(d)$ is a circularly
symmetric, 2D-Gaussian which is a function of distance $d$ from
SNr$_j$ [Fig.~\ref{fig:ccg_model}(c)]. A similar scheme applied for
the connectivity from Str\_D2 to GPe and for a number of the other
connections in the \ccg~model; in the SpineML implementation of the
model, this connectivity scheme is named `GaussianKernel'.
% NB - having this Fig. ref here is important for figure ordering!
Fig.~\ref{fig:brainmodel} shows a schematic of the SpineML implementation
of the model, based on a diagram output directly from
SpineCreator. Populations for Str\_D1, Str\_D2, STN, SNr and GPe are
shown within the `Basal Ganglia' box. Input comes into the model via
the `World' population and the output population is SC\_deep. Compare
this diagram with Figs.~\ref{fig:ccg_model}(a) \&
(b). Fig.~\ref{fig:brainmodel} expands the `SC', `BG' and `slow retinal'
boxes from Fig.~\ref{fig:ccg_model}(a).}

% Now some FEF stuff taken from Alex's pre-pub
\rvwt{The frontal eye fields (FEF) are a key cortical area for the
generation of saccadic eye movements
\citep{hikosaka_role_2000,tehovnik_eye_2000,robinson_eye_1969,bruce_primate_1985}.
Saccadic targets are mapped retinotopically over its
surface \citep{robinson_eye_1969,bruce_primate_1985,sabes_parietal_2002},
and increased neural activity at a location in the map precedes a
saccade to that location. Importantly, the FEF is also associated with
visual decision making
\citep{thompson_visual_2005,schall_saccade_1995,monosov_measurements_2008,cohen_neural_2009}.
Thus, in a saccade choice, increased FEF activity is predictive of the
eye movement whether correct or
incorrect \citep{thompson_frontal_2005}, rather than of the correct
response.}

\rvwt{FEF neurons can be divided in to three functional groups,
related to whether their activity corresponds with visual stimuli,
motor action, or both \citep{segraves_functional_1987}. The \ccg~model
simplifies this categorisation using a single layer of 50 by 50 units
representing the mean of all three groups. This layer therefore
responds to both visual stimuli and the buildup of activity associated
with motor (saccadic) action. The retina provides a persistent
luminance signal into the FEF through the dorsal visual
pathway \citep{ungerleider_two_1982} which is abbreviated in this
model to a direct connection with delay.}

@@

\rvwt{The superior colliculus (SC) is a sub-cortical nucleus which
also plays a critical r\^ole in the generation of
saccades \citep{hikosaka_visual_1983}. Both FEF and SC have direct
connections to the saccadic burst generator (SBG, see
Sect.~\ref{sec:methods:brainstem}). If either is lesioned, the other
can direct gaze, following a period of
adjustement \citep{latto_effects_1977}, albeit with some persistent
deficits. The SC is also a direct target of output from the
SNr \citep{jayaraman_nigrotectal_1977,jiang_opposing_2003} and can be
influenced by the action selection mechanisms of the BG. In
particular, it forms a loop with BG, but unlike its cortical
counterpart in FEF, the input to basal ganglia comes via the thalamus
[Fig~\ref{fig:ccg_model}(a), blue arrows].}

\rvwt{While the SC has seven alternating cell-~and fibre-rich
layers \citep{wurtz_visual-motor_1980}, in most cases these are
divided into the `superficial' and `deep' layers, which have
significantly different response properties. Cells in the superficial
layers, which receive input from the retina, are mainly visually
responsive, with a preferred response to phasic events (luminance
onsets and offsets) and movement on the visual
field \citep{goldberg_activity_1972}. In contrast, cells in the deep
layers receive multi-modal input, including inhibitory input from the
output structures of the BG \citep{jayaraman_nigrotectal_1977}, and
are directly involved in the generation of saccadic eye
movements. Saccade related activity in the deep layers appears to
generate saccades through `population coding', with a weighted sum of
activity across the retinotopy of SC determining the saccade target
\citep{lee_population_1988,van_opstal_role_1990,mays_dissociation_1980}.
The deep layers of SC receive input from the FEF in a topographic
manner \citep{stanton_frontal_1988,sommer_composition_2000}.}

@@

% Foveal mask
\rvwt{The \ccg~model incorporates a special connectivity pattern for
visual input via cortical (FEF) and sub-cortical (thalamus)
pathways. Due to the retinotopic mapping (Sect.~\ref{sec:mapping}),
foveal luminances deliver a strong signal to the BG; roughly one third
of the map is activated for the foveal targets used in this work
(Fig.~\ref{fig:mapping}, red cross). This makes it virtually impossible
for a peripheral target (Fig.~\ref{fig:mapping}, yellow cross) to win
selection in the BG. Even if the peripheral target competed
successfully to generate a saccade, this process would cause a
significant delay, leading to latencies much larger than those
observed experimentally. To overcome this problem, the \ccg~model
incorporates a mechanism in which the synaptic strength of connections
between FEF, thalamus and striatum are reduced close to the fovea
according to a shifted hyperbolic tangent. This connection is named
`DecayingAtFovea' in the SpineML implementation and follows a modified
sigmoidal curve rather than tanh. In either case, the relation is
`S-shaped' and normalised to the range [0 1]. Far from the fovea
(where the S-shaped curve has the value $\approx$1), the connectivity
pattern looks almost identical to a one-to-one connection.}

@@

% KM
\rvwt{An illustration of the biomechanical eye model is given in
Figure \ref{fig:bioeye}, while Figure \ref{fig:bioeye} depicts
the head model used in the proposed framework.}

>>>>

% Solution described means we now need to do background of eye movements:
In an attempt to build a model combining brain, realistic
biomechanics \e{and} sensory feedback, we sought to extend our
\rvw{previous work modelling the oculomotor system by adding a virtual,
biomechanical eye model able to make physically realistic
movements. The rotational state of the eye would then determine how
visual features in the virtual world were projected back into the
brain model.}
%
The existing
model \citep{cope_basal_2017} is already able to capture sensory input
and convert it into a neural signal, assumed to specify the target of
a \e{saccadic eye movement}; a fast movement of the eyes which directs
the fovea to a region of interest in the field of view.
%
The oculomotor system is an excellent candidate for modelling because
its movements can be specified with only three degrees of freedom,
making it one of the simplest neuro-muscular systems in the body.
%
It is nevertheless behaviourally interesting, as saccadic eye
movements reveal information about decision making at a subconscious
level
\citep{deubel_saccade_1996,reppert_modulation_2015,marcos_determining_2016}.
The modelling of the oculomotor system is served by a large body of
behavioural data describing
saccades \citep{tipper_reaching_2001,walker_effect_1997,casteau_effect_2012},
many anatomical studies of the neural substrates involved
\citep{meredith_intrinsic_1998,isa_intrinsic_2002,isa_exploring_2009}
and electrophysiological data linking these together
\citep{hepp_spatio-temporal_1983,dorris_neuronal_1997,mcpeek_competition_2003,vokoun_circuit_2011}.
Furthermore, in the context of building \e{behaving} systems, a
realistic mechanism for gathering visual information is a necessary
part of any model for which the behaviour requires visual attention
and decision making.  This is obvious from extrinsic
considerations---a subject must look at a scene to make decisions or
navigate within it. It also follows for \e{intrinsic} reasons. For
example, \cite{howard_hand_1997} showed that visual cues affect reach
trajectories and the same group later demonstrated that reaching
affects the saccadic system
\citep{tipper_reaching_2001} suggesting a close relationship between these
neural systems. \rvw{Building a behaving oculomotor system will
therefore assist future computational neurobehavioural modelling
efforts that involve reaching.}

Many neural populations are involved in the coding of saccadic eye
movements, only a very brief overview is given here; for a review,
see \cite{munoz_commentary:_2002}. One pathway takes information from
the retina directly into the superficial layers of the superior
colliculus in the brainstem
\citep{sterling_receptive_1971,linden_massive_1983,wu_involvement_1994}.
Activity within the superior colliculus then excites neurons in the
pons, medulla and rostral mid-brain \citep{sparks_brainstem_2002}
and \rvw{finally the motor neurons}, which innervate the extraocular
muscles
\citep{fuchs_firing_1970,sparks_brainstem_2002}. This direct pathway
is responsible for the low latency saccades called express saccades
\citep{schiller_effect_1987,edelman_activity_1996}.
%
Information from the retina is also processed by visual cortex which
feeds through to the frontal eye fields in which activity is related
to reflexive and voluntary
saccades \citep{schall_neural_1999}. Activity build-up in the frontal
eye fields is transferred to the intermediate layers of the superior
colliculus
\citep{stanton_frontal_1988-1} and is also processed by the basal ganglia,
which participates in the selection of the winning saccade end point
\citep{stanton_frontal_1988,hikosaka_role_2000}.
%
Although both cortical and subcortical paths produce a saccade target
signal in the superior colliculus, it is also possible for animals to
make relatively normal saccades even after the colliculus has been
ablated
\citep{wurtz_activity_1972,aizawa_reversible_1998},
though express saccades are lost with collicular lesions
\citep{schiller_effect_1987}. This makes the superior colliculus
a perplexing structure, being both critically involved in saccade
target specification \citep{sparks_sensory_1987} and saccade dynamic control
\citep{waitzman_superior_1991,goossens_optimal_2012} and yet dispensible.
The `backup pathway' likely incorporates the oculomotor vermis and
fastigial oculomotor region of the cerebellum which are known to
participate in the specification, dynamics and adaptation of saccadic
eye movements \citep{kleine_saccade-related_2003,takagi_effects_1998}.

@@

% A very short introduction to the section.
The integrated brain and biomechanical model described here is a
development of the model in \cite{cope_basal_2017}, referred to here
as the \e{\ccg~model}. This was a rate-coded neural network model
incorporating retinal populations, frontal eye fields (FEF), the basal
ganglia (BG), and the superior colliculus (SC).  The \ccg~model takes
as \e{input} the positions of luminances \rvwt{(of fixed shape and
intensity)} on a topographic map. \rvwt{While certain assumptions
were made about the input---that a luminant input excites activity on
a retinotopic layer, with computer code carrying out the
transformation achieved in the brain by a neural connectivity
map \citep{thivierge_topographic_2007}---it is nonetheless \e{not} an
input-assumption model according to our definition because the
activity generated in the neural input layer is modelled as a response
to the luminances, rather than being crafted.}  In the \ccg~model, the
centroid of the activity in the deep layers of superior colliculus was
assumed to accurately encode the location of the eye at the end of the
saccade
\citep{wurtz_activity_1972,robinson_eye_1972,van_gisbergen_collicular_1987,mcilwain_lateral_1982}.  This location was used to
recalculate the positions of the luminances in the eye's frame of
reference at each time step. \rvwt{Because a pattern of neural
activity in the output population was assumed to have a behavioural
outcome, it was thus an \e{output-assumption model}}.  The model
included no brainstem populations other than superior colliculus, nor
a neuromuscular model.

To the \ccg~model, we added a brainstem model and a biomechanical eye
model. \rvw{The rate-coded brainstem model was taken from the
literature} \citep{gancarz_neural_1998} as the best-of-breed saccadic
burst generator \citep{girard_brainstem_2005}.  The biomechanical eye
was implemented using the biomechanical modelling framework OpenSim;
the brain and brainstem were modelled using the SpineML toolchain.
These will be described below, \rvwt{along with a review of
the \ccg~model}, but first we will give a description of the
co-ordinate systems that were used.

@@

% Describe what is modelled first. This para is 'The OM system
% performs action selection and the BG will do that for us.'
The brain model, excluding the brainstem, is a re-implementation of
the \ccg~model \rvwt{of reflexive saccadic behaviour \citep{cope_basal_2017}}.
\rvwt{Reflexive saccades are fast eye movements elicited by abrupt
changes in the peripheral visual scene (reflexive saccades can occur
also as a result of auditory and somatosensory stimuli, but these
modalities are ignored in this model). A reflexive saccade has a
starting position defined by the initial orientation of the eye and an
end-point position in which the eye is directed towards a new
target. Regardless of the number of targets within the visual scene,
the brain must choose one location as the end-point, because the eyes
can look only in one direction at a time. The functionality reproduced
by the \ccg~model is `the selection of the best target end-point for a
reflexive saccade'. A competition such as this between incompatible
movements is often referred to as an \e{action selection}
problem \citep{norman_attention_1986,maes_dynamics_1989,redgrave_basal_1999}.
The \ccg~model is therefore a model of action-selection in the
oculomotor system for reflexive saccades.  One hypothesis for the
r\^ole played by the basal ganglia (BG) is that the system
performs \e{action selection}
\citep{mink_basal_1996,redgrave_basal_1999,hikosaka_role_2000}.
The \ccg~model places the BG at the centre of the oculomotor system;
this follows the known anatomy of the
region \citep{hikosaka_role_2000} and provides a mechanism for action
selection of the best saccade. The BG receives input indirectly from
the superior colliculus, which has a retinotopic
arrangement \citep{ottes_visuomotor_1986}.}

@@

% Now describe each bit a little more, with help of the CCG
% macroscopic architecture figure.
\rvwt{Fig.~\ref{fig:ccg_model}(a) shows the macroscopic architecture of
the \ccg~model. The figure shows the relationships between the retinal
input populations, the FEF, the populations comprising the BG
sub-system (the red border indicates that the box represents a number
of populations as a sub-system), the thalamus and the superior
colliculus. Excitatory connections are indicated with arrowheads;
inhibitory connections with circles in place of the arrowheads. The
blue and green connection lines indicate two thalamo-basal ganglia
loops, one cortical loop through FEF (green), the other a sub-cortical
loop through SC. It is important to note that although they are given
different colours in the diagram, these loops are in no way
independent, with loop activity combining both in thalamus and in the
basal ganglia and a direct excitatory, feed-forward connection from
FEF to SC.}

@@

% Dopamine, tonic output.
\rvwt{At the neuronal level, the STN, GPe and SNr have tonic output
activity \citep{chevalier_disinhibition_1990,delong_primate_1985,kita_intracellular_1991}. This
is modelled using piecewise linear output functions with zero offsets,
$c$ (see Eq.~\ref{eq:lintransfunc}) but with noise added to the
input. In striatum, Str\_D1 and Str\_D2 have positive offset $c$,
mimicking the so-called `down-state' of medium spiny neurons which
have a resting potential far below spiking threshold and require
co-ordinated input to generate action
potentials \citep{wilson_origins_1996}.
% Note, in Alex's paper, there is an error with the offsets for D1,
% D2. These are set to -0.05 instead of 0.05. A quick trial with the
% model suggests this doesn't make a very large difference to the
% overall operation (you just see a background of noise around the
% signals in Str), and I have the sign correct in any case.
In addition, the Str\_D1 and Str\_D2 neurons are influenced by
dopamine in different ways; facilitating cortico-striatal transmission
at medium spiny neurons with D1
receptors \citep{hernandez-lopez_D1_1997,gonon_prolonged_1997} and
reducing transmission at those with D2
receptors \citep{delgado_inhibitory_1999}. These effects are modelled
using a dopamine parameter, $d$, which modulates the input activations
$a_{in}^{D1}$ and $a_{in}^{D2}$ as:
\begin{equation}
   a_{in}^{D1} = (0.2 + d)A
\end{equation}
\begin{equation}
   a_{in}^{D2} = (1 - d)A
\end{equation}
where $A$ is the input activation (see also
Eqns.~\ref{eq:a_in_d1} \& \ref{eq:a_in_d2}). For the `normal, healthy'
value for $d$ of 0.7, Str\_D1 activation is relatively enhanced
(0.9$A$); Str\_D2 activation is one third of this value (0.3$A$). The
major effect of this difference in the relative strength of the
activity in Str\_D1 versus Str\_D2 is simply that a change in the
level of activity in Str\_D1 affects the off-centre, on-surround
mechanism. The effect of varying the input into Str\_D2 is much more
subtle, with only a small change in the amount of inhibition fed from
GPe into STN (via a focussed, one-to-one connection) being affected by
the change, along with a small change in the inhibition fed into SNr
from GPe (also via a one-to-one connection). It is not possible to
ascribe the dynamic effect of the dopamine parameter to any single
population, because the activity is recurrently
connected through multiple loops. Thus, a line of reasoning such as
``reduced activity for a luminance in Str\_D2 will lead to less
inhibition in that region in GPe, which means that there will be
higher activity there, and hence more inhibition for that region
passed to STN leading us to expect less activity in STN'' is verified
by running a suitable simulation with the model, but the effect is
small. }
%
\rvwt{Note that
the effect of dopamine in the model is only to modulate the strength
of cortico-striatal synapses; no learning is modelled and so the
significance of dopamine as a prediction error signal is outside the
scope of the current work.}

% Change this - we already talked about channels in the context of the
% three channel model shown. Now talk about how we have a map of
% channels - a continuous space in the \ccg model. This can then lead
% on to Fig 3.
\rvwt{The GPR model in Fig.~\ref{fig:ccg_model}(b) has only three
channels, with the focussed inhibition from striatum to SNr and GPe
defined by a simple one-to-one scheme. The action channels represent
discrete, incompatible motor action choices. In the oculomotor model,
an action channel represents the end-point of saccade, and the
competition carried out in the basal ganglia is between potential
saccade end-points. However, eye movements have a \e{continuous}
end-point space; the eye can rotate to any orientation within its
biomechanically permissible range. Some end-points within this range
are mutually exclusive---it's not possible to look to the left and to
the right simultaneously---but \e{nearby} end-points are not
necessarily incompatible. A small enough error in the end-point of a
saccade will not prevent the eye from foveating on a target as the
foveal region of high visual acuity is not infinitesimally small. To
cope with this requirement, the populations within the oculomotor
basal ganglia are conceived of as two-dimensional topographic grids of
leaky integrator neural elements. Activity in each element corresponds
to a spatial location in the visual field. Neighbouring elements
correspond to locations which are close to each other in the visual
field. Focussed one-to-one projections in the GPR model are replaced
by projective fields with many weighted connections. Specifically,
each unit in Str\_D1 projects to a counterpart SNr$_j$ in SNr with
some weight $w_{max}$, but also connects to neighbouring nodes in SNr
with a weight given by $w_{max}.G(d)$, where $G(d)$ is a circularly
symmetric, 2D-Gaussian which is a function of distance $d$ from
SNr$_j$ [Fig.~\ref{fig:ccg_model}(c)]. A similar scheme applies for
the connectivity from Str\_D2 to GPe and for a number of the other
connections in the \ccg~model; in the SpineML implementation of the
model, this connectivity scheme is named `GaussianKernel'.
% NB - having this Fig. ref here is important for figure ordering!
Fig.~\ref{fig:brainmodel} shows a schematic of the SpineML implementation
of the model, based on a diagram produced by SpineCreator. Populations
for Str\_D1, Str\_D2, STN, SNr and GPe are
shown within the `Basal Ganglia' box. Input comes into the model via
the `World' population and the output population is SC\_deep. Compare
this diagram with Figs.~\ref{fig:ccg_model}(a) \&
(b). Fig.~\ref{fig:brainmodel} expands the `SC', `BG' and `slow retinal'
boxes from Fig.~\ref{fig:ccg_model}(a).}

% Now some FEF stuff taken from Alex's pre-pub
\rvwt{The frontal eye fields (FEF) are a key cortical area for the
generation of saccadic eye movements
\citep{hikosaka_role_2000,tehovnik_eye_2000,robinson_eye_1969,bruce_primate_1985}.
Saccadic targets are retinotopically mapped on its
surface \citep{robinson_eye_1969,bruce_primate_1985,sabes_parietal_2002},
and increased neural activity at a location on the map precedes a
saccade to that location. Importantly, the FEF is also associated with
visual decision making
\citep{thompson_visual_2005,schall_saccade_1995,monosov_measurements_2008,cohen_neural_2009}.
Thus, in a saccade choice, increased FEF activity is predictive of the
eye movement whether correct or
incorrect \citep{thompson_frontal_2005}, rather than of the correct
response.}

\rvwt{FEF neurons can be divided into three functional groups,
related to whether their activity corresponds with visual stimuli,
motor action, or both \citep{segraves_functional_1987}. The \ccg~model
simplifies this categorisation using a single layer of 50 by 50 units
representing the mean of all three groups. This layer therefore
responds to both visual stimuli and the buildup of activity associated
with motor (saccadic) action. The retina provides a persistent
luminance signal into the FEF through the dorsal visual
pathway \citep{ungerleider_two_1982} which is abbreviated in this
model to a direct connection with delay and noise.}

@@

\rvwt{The superior colliculus (SC) is a sub-cortical nucleus which
also plays a critical r\^ole in the generation of
saccades \citep{hikosaka_visual_1983}. Both FEF and SC have direct
connections to the saccadic burst generator (SBG, see
Sect.~\ref{sec:methods:brainstem}). If either is lesioned, the other
can direct gaze, following a period of
adjustment \citep{latto_effects_1977}, albeit with some persistent
deficits. The SC is also a direct target of output from the
SNr \citep{jayaraman_nigrotectal_1977,jiang_opposing_2003} and can be
influenced by the action selection mechanisms of the BG. In
particular, it forms a loop with BG, but unlike its cortical
counterpart in FEF, the input to basal ganglia comes via the thalamus
[Fig~\ref{fig:ccg_model}(a), blue arrows].}

\rvwt{While the SC has seven alternating cell-~and fibre-rich
layers \citep{wurtz_visual-motor_1980}, in most cases these are
categorised as the `superficial' and `deep' layers, which have
significantly different response properties. Cells in the superficial
layers, which receive input from the retina, are mainly visually
responsive, with a preferred response to phasic events (luminance
onsets and offsets) and movement on the visual
field \citep{goldberg_activity_1972}. In contrast, cells in the deep
layers receive multi-modal input, including inhibitory input from the
output structures of the BG \citep{jayaraman_nigrotectal_1977}, and
are directly involved in the generation of saccadic eye
movements. Saccade related activity in the deep layers appears to
generate saccades through `population coding', with a weighted sum of
activity across the retinotopy of SC determining the saccade target
\citep{lee_population_1988,van_opstal_role_1990,mays_dissociation_1980}.
The deep layers of SC receive input from the FEF in a topographic
manner \citep{stanton_frontal_1988,sommer_composition_2000}.}

@@

% Foveal mask
\rvwt{The \ccg~model incorporates a special connectivity pattern for
visual input to the BG via cortical (FEF) and sub-cortical (thalamus)
pathways. Due to the retinotopic mapping (Sect.~\ref{sec:mapping}),
foveal luminances deliver a strong signal to the BG; roughly one third
of the map is activated for the foveal targets used in this work
(Fig.~\ref{fig:mapping}, red cross). This makes it virtually impossible
for a peripheral target (Fig.~\ref{fig:mapping}, yellow cross) to win
selection in the BG. Even if the peripheral target competed
successfully to generate a saccade, this process would cause a
significant delay, leading to latencies much larger than those
observed experimentally. To overcome this problem, the \ccg~model
incorporates a mechanism in which the synaptic strength of connections
between FEF, thalamus and striatum are reduced close to the fovea
according to a shifted hyperbolic tangent. This connection is named
`DecayingAtFovea' in the SpineML implementation and follows a modified
sigmoidal curve rather than tanh. In either case, the relation is
`S-shaped' and normalised to the range [0 1]. Far from the fovea
(where the S-shaped curve has the value $\approx$1), the connectivity
pattern looks almost identical to a one-to-one connection.}

@@

% KM
\rvwt{An illustration of the biomechanical eye model is given in
Figure \ref{fig:bioeye}(a), while Figure \ref{fig:bioeye}(b) depicts
the head model used in the proposed framework.}
